Who does big team science?

### 2023-03-02 manuscript_scopus.Rmd
rearranging for clarity
adding pictures for sav

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
output            : papaja::apa6_docx
---

```{r setup, include = FALSE}
library(papaja)
library(knitr)
```

# Significance

@@

The credibility movement has been mirrored by the calls for
diversification or de-WEIRDing (e.g., Western, Educated, Industrialized,
Rich, and Democratic) scientific research [@henrich2010; @newson2021;
@rad2018] by improving representation in research samples. Like the
large-scale studies in Physics [@aphilos2021; @castelnovo2018] and
Biology [@collins2003], the social sciences struggle to represent the
breadth of humanity across both researcher and population
characteristics. Now, grassroots organizations, such as the
Psychological Science Accelerator [@moshontz2018], ManyBabies
(<https://manybabies.github.io/>), NutNet (<https://nutnet.org/>), and
DRAGNet (<https://dragnetglobal.weebly.com/>) can begin to tackle these
issues by recruiting research labs from all over the globe to provide
diversity in geographic, linguistic, and researcher representation.
Publications have examined the global understanding of morality, face
processing, COVID-19 information signaling, and more [@dorison2022;
@wang2021; @legate2022; @bago2022; @jones2021; @vanbavel2022]. While
these organizations and one-time groups for BTS studies have provided an
incredible wealth of data for the scientific community, we do not yet
know exactly *who* is involved with, and benefits from, the BTS and
credibility movement. Publications on BTS generally explore challenges,
lessons learned, the need for BTS [@forscher2020; @coles2022].

@@

# Approach

@@

We have defined **BTS publications** as publications with at least 10
authors at 10 different institutions that were published in
peer-reviewed journals or had posted a full paper pre-print. We will use
data from 1970 and forward in the publications (`ani)` database, as it
is noted online that this time period includes cited references for
calculation of several of our variables described below. We will analyze
our results based on the big four subject areas: Physical Sciences,
Health Sciences, Social Sciences, and Life Sciences.

@@

Using these criteria, we will extract the following information for
publication sources: the name of the publication (source title), subject
area (both the large four subject areas and the smaller four digit all
science journal classification ASJC codes). We will examine journal
impact using the Source Normalized Impact per Paper from the journal
(`sources)` database.

@@

-   We will present visualizations of the country information listed for
    authors, and we will discuss the areas of world in which authors
    generally come from, as well as the lowest representation of
    authors.
-   To understand the change in representation diversity, we will
    summarize the total number of geopolitical regions for each paper.
    Using a linear model, we will examine if the number of regions
    present is predicted by the year of publication. Increasing
    diversity would be represented by a positive slope, while decreasing
    diversity would be represented by a negative slope.
-   Last, we will examine the differences in representation for
    corresponding author sets versus all other authors. For papers with
    10 to 49 authors, we will use the three first authors and the last
    author to compare against other authors. For 50 to 99 authors, five
    first authors plus last will be used, and for all papers with more
    than 100 authors, we will use ten first authors and the last author.
    We will calculate the frequencies of each of the UN Sub-Regions for
    first authors versus other authors, converting these values to
    proportions. Given the expected small sample sizes of these
    contingency tables, we will group together titles based on the year
    of publication (assuming at least 5 publications per year, these may
    be binned by 5-year or smaller increments to increase sample size).
    For each grouping, we will calculate the effect size of the
    differences in frequencies comparing first authors to all other
    authors. Since this data is categorical, we will use Cramer's V to
    represent the effect size. If the effect size includes zero in its
    confidence interval, this result will imply that first and all other
    authors represent the same pattern of UN Sub-Region diversity. Any
    confidence interval that does include zero represents a difference
    in diversity. We will report these values and discuss what regions
    of the world are represented when effect sizes indicate a different
    from zero using standardized residuals.

>>>>

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja'`)"
documentclass     : "apa7"
output            : papaja::apa6_docx
---

```{r setup, include = FALSE}
#knitr::opts_knit(echo = TRUE)
library(papaja)
library(knitr)
library(ggplot2)
library(treemapify)
library(countrycode)
library(rio)
library(maps)
library(dplyr)
```

```{r data, include = FALSE}
geo_country <- import("py_outputs/geo_country.xlsx")
```

@@

The credibility movement has been mirrored by the calls for
diversification or de-WEIRDing (e.g., Western, Educated, Industrialized,
Rich, and Democratic) scientific research [@henrich2010; @newson2021;
@rad2018] by improving representation in research samples. Like the
large-scale studies in Physics [@aphilos2021; @castelnovo2018] and
Biology [@collins2003], the social sciences struggle to represent the
breadth of humanity across both researcher and population
characteristics. Now, grassroots organizations, such as the
Psychological Science Accelerator [@moshontz2018], ManyBabies
(<https://manybabies.github.io/>), NutNet (<https://nutnet.org/>), and
DRAGNet (<https://dragnetglobal.weebly.com/>) can begin to tackle these
issues by recruiting research labs from all over the globe to provide
diversity in geographic, linguistic, and researcher representation.
Publications have examined the global understanding of morality, face
processing, COVID-19 information signaling, and more [@dorison2022;
@wang2021; @legate2022; @bago2022; @jones2021; @vanbavel2022]. While
these organizations and one-time groups for BTS studies have provided an
incredible wealth of data for the scientific community, we do not yet
know exactly *who* is involved with, and benefits from, the BTS and
credibility movement. Publications on BTS generally explore challenges,
lessons learned, and the need for BTS [@forscher2020; @coles2022].

@@

# Research Questions

@@

We have defined **BTS publications** as publications with at least 10
authors at 10 different institutions that were published in
peer-reviewed journals or had posted a full paper pre-print. We will use
data from 1970 and forward in the publications (`ani`) database, as it
is noted online that this time period includes cited references for
calculation of several of our variables described below. We will analyze
our results based on the big four subject areas: Physical Sciences,
Health Sciences, Social Sciences, and Life Sciences.

@@

Using these criteria, we will extract the following information for
publication sources: the name of the publication (source title), subject
area (both the large four subject areas and the smaller four digit all
science journal classification ASJC codes). We will examine journal
impact using the Source Normalized Impact per Paper from the journal
(`sources`) database.

@@

-   We will present visualizations of the country information listed for
    authors, and we will discuss the areas of world in which authors
    generally come from, as well as the lowest representation of
    authors.

```{r geo-map}
# serbia has changed codes
geo_country$affiliation_tag_country <- toupper(geo_country$affiliation_tag_country)
geo_country <- subset(geo_country, affiliation_tag_country != "NONE")
geo_country$affiliation_tag_country[geo_country$affiliation_tag_country == "SCG"] <- "SRB"

geo_country <- geo_country %>% 
  group_by(affiliation_tag_country) %>% 
  summarize(count = sum(count))

# convert country code to region code
# this will warn you about the ones that have multiple countries 
# create a world map 
world_map <- map_data(map = "world")
world_map$orig_region <- world_map$region
world_map$region <- iso.alpha(world_map$region, n = 3)
world_map <- subset(world_map, region != "ATA")

# make a map on a continuous scale
ggplot(geo_country) +
  geom_map(aes(map_id = affiliation_tag_country, fill = count), map = world_map) +
  geom_polygon(data = world_map, 
               aes(x = long, y = lat, group = group), 
               colour = 'black', fill = NA) + 
  theme_void() + 
  scale_fill_distiller(name = "Sample Size",
                       palette = "Greys",
                       direction = 1,
                       na.value = "white")

ggsave("figure/continuous_country.png")

# maybe try binning
geo_country$count_binned <- if_else(
  geo_country$count >= 10000, "10,000+", 
  if_else(
    geo_country$count < 10000 & geo_country$count >= 5000, "5,000-9,999",
    if_else(
      geo_country$count < 5000 & geo_country$count >= 1000, "1,000-4,999",
      if_else(
        geo_country$count < 1000 & geo_country$count >= 100, "100-999",
        "< 100"
      )
    )
  )
)

geo_country$count_binned <- factor(geo_country$count_binned, 
                                   levels = unique(geo_country$count_binned))

# map of binned data 
bin_country <- ggplot(geo_country) +
  geom_map(aes(map_id = affiliation_tag_country, fill = count_binned), map = world_map) +
  geom_polygon(data = world_map, 
               aes(x = long, y = lat, group = group), 
               colour = 'black', fill = NA) + 
  theme_void() + 
  scale_fill_manual(name = "Sample Size",
                    values = c( "black", "#323232", "#646464","#969696","#c8c8c8"))

bin_country

ggsave("figure/binned_country.png", bin_country)

# tree map
geo_country$un_region_sub <- countrycode(
  sourcevar = geo_country$affiliation_tag_country,
  origin = 'iso3c', 
  destination = 'un.regionsub.name'
)

geo_country$un_region <- countrycode(
  sourcevar = geo_country$affiliation_tag_country,
  origin = 'iso3c', 
  destination = 'un.region.name'
)

geo_country$un_region[geo_country$affiliation_tag_country == "TWN"] <- "Asia"
geo_country$un_region_sub[geo_country$affiliation_tag_country == "TWN"] <- "Eastern Asia"

geo_country <- geo_country %>% 
  na.omit()

tree <- ggplot(geo_country, aes(area = count, fill = count_binned,
               label = affiliation_tag_country, subgroup = un_region_sub)) +
  geom_treemap() +
  geom_treemap_subgroup_border(colour = "white", size = 5) +
  # geom_treemap_subgroup_text(place = "top", grow = TRUE,
  #                            alpha = 0.25, colour = "black",
  #                            fontface = "italic") +
  geom_treemap_text(colour = "white", place = "centre",
                    size = 15, grow = FALSE) +  
  scale_fill_manual(name = "Sample Size",
                    values = c("#c8c8c8", "#969696", "#646464", "#323232", "black")) 
  # scale_fill_gradient(name = "Sample Size",
  #                     low = "#c8c8c8", 
  #                     high = "#323232")

tree

ggsave("figure/treemap.png",
       tree)
```

-   To understand the change in representation diversity, we will
    summarize the total number of geopolitical regions for each paper.
    Using a linear model, we will examine if the number of regions
    present is predicted by the year of publication. Increasing
    diversity would be represented by a positive slope, while decreasing
    diversity would be represented by a negative slope.

-   Last, we will examine the differences in representation for
    corresponding author sets versus all other authors. For papers with
    10 to 49 authors, we will use the three first authors and the last
    author to compare against other authors. For 50 to 99 authors, five
    first authors plus last will be used, and for all papers with more
    than 100 authors, we will use ten first authors and the last author.
    We will calculate the frequencies of each of the UN Sub-Regions for
    first authors versus other authors, converting these values to
    proportions. Given the expected small sample sizes of these
    contingency tables, we will group together titles based on the year
    of publication (assuming at least 5 publications per year, these may
    be binned by 5-year or smaller increments to increase sample size).
    For each grouping, we will calculate the effect size of the
    differences in frequencies comparing first authors to all other
    authors. Since this data is categorical, we will use Cramer's V to
    represent the effect size. If the effect size includes zero in its
    confidence interval, this result will imply that first and all other
    authors represent the same pattern of UN Sub-Region diversity. Any
    confidence interval that does include zero represents a difference
    in diversity. We will report these values and discuss what regions
    of the world are represented when effect sizes indicate a different
    from zero using standardized residuals.
